---
title: "Anomaly Detection"
author: "Mercy M. Kubania"
date: "7/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem Definition
> As a Data analyst at Carrefour Kenya I am doing a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). 

## a) Specifying the Question

1. Identify Anomalies in the Sales Data

## b) Metric of Success
> Able to detect Fraud

## c) The Context
Working on the CarreFour Kenya Sales data to identify Fraud in their Sales Data

## d) Experimental Design
1. Source the Data
2. Load the Data
3. Check the Data and Clean it
4. Perform Exploratory Data Analysis
5. Implement the Solution
6. Challenge the Solution

# Data Sourcing
```{r Anomaly Data}
# Data to Detect Anomaly
sales <- read.csv("~/Documents/R markdowns/IP Week 14/Supermarket_Sales_Forecasting - Sales.csv")

head(sales)
```


# Check the Data
```{r Structure}
str(sales)

```
```{r Missing Data}
anyNA(sales)
colSums(is.na(sales))

```

```{r Convert to a Tibble, message=TRUE}
# Convert the data into Tibble and

library(tidyverse)
library(tibbletime)

sales$Date <- as.Date(sales$Date, "%m/%d/%Y")

sales_ts = sales %>%
  as.tibble() %>%
  as_tbl_time(Date) %>%
  arrange(Date) %>%
  as_period("daily")

```

# Data Cleaning
```{r Missing Values}
anyNA(sales_ts)

```


# Implement the Solution
```{r Anomaly Detection}
library(anomalize)

sales_anomaly <- sales_ts %>%
  time_decompose(Sales) %>% 
  anomalize(remainder,max_anoms = 0.2, alpha=0.05) %>%
  time_recompose() %>% glimpse()

# Plot
sales_anomaly %>% plot_anomalies(time_recomposed = TRUE)
#sales_anomaly %>% plot_anomaly_decomposition(time_recompose = T)

```


There is are no Anomalies in our Sales Data 

