---
title: "Assosiation Analysis for CarreFour Kenya"
author: "Mercy M. Kubania"
date: "7/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Definition
> As a Data analyst at Carrefour Kenya I am doing a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). 

## a) Specifying the Question

1. Identify relationships between the Group of Items

## b) Metric of Success
> Identify the relationships between the Groups of Items provided

## c) The Context
Working on the CarreFour Kenya Sales data to identify the most relevant marketing strategies by identifying related groups of Items that are related

## d) Experimental Design
1. Source the Data
2. Load the Data
3. Check the Data
4. Implement the Solution
5. Challenge the Solution

# Data Sourcing
```{r Library}
library(arules)
```

```{r Association Data}
# Data to find associations
transaction <- read.transactions("~/Documents/R markdowns/IP Week 14/Supermarket_Sales_Dataset II.csv",sep=",")
transaction

```
# Check the Data
```{r Check Data Type}
# Confirm that our data type is transaction
class(transaction)

```

```{r Inspect}
# The First 10 transactions
inspect(transaction[1:10])

```
```{r Items Set size}

# number of items in the first 10 itemset or transaction
size(head(transaction, 10))

```

```{r Items}
# View 10 items that make up our dataset

item <- as.data.frame(itemLabels(transaction))
colnames(item) <- "Item"
head(item,10)
```

```{r Summary}
# Summary of our Transactions
summary(transaction)

```
```{r Explore Item frequency}
print("Frequency of each item")
itemFrequency(transaction[,1:15], type="absolute")

cat ("Tab\n")

print("Frequency of first 15 Items compared to Others")
# How many times these items have appeared as compared to others.
round(itemFrequency(transaction[,1:15], type="relative")*100,2)


```
```{r Item Frequency Plot}
# Most common items
library(RColorBrewer)

itemFrequencyPlot(transaction, topN= 10, col=brewer.pal(8,'Set1'),main="Relative Item Frequency Plot")

itemFrequencyPlot(transaction, support= 0.1, col=brewer.pal(8,'Set1'),main="Relative Importance is at least 10%")

```


# Implement the Solution
```{r Rules}
# Build a Model based on association rules
# min_supp -> 0.001
# confidence -> 0.8

rules <- apriori(transaction, parameter = list(supp=0.001, conf=0.75))
rules

summary(rules)

inspect(rules[1:5])
```
```{r }
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:10])
```

# Challenge the Solution