---
title: "Marketing Strategies for Carrefour Kenya"
author: "Mercy M. Kubania"
date: "7/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Statement
> As a Data analyst at Carrefour Kenya I am doing a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). 

## a)Specifying the Question
1. Build models that will inform on relevant marketing strategies that will result in the highest no. of sales (total price including tax) for the Marketing Department at CarreFour Kenya

## b)The Metric of Success
> Analyse the data and Build a model that will inform on the relevant marketing strategies for the Marketing Department

## c)The Context
> Working on the CarreFour Kenya Sales data to identify the most relevant marketing strategies


## d)Experimental Design
1. Source the Data
2. Load the Data
3. Check the Data and Clean it
4. Perform Exploratory Data Analysis
5. Implement the Solution
6. Challenge the Solution

# Data Sourcing
```{r Dimension and Feature Data}
# Data for Analysis, Dimension Reduction and Feature Selection

sales <- read.csv("~/Documents/R markdowns/IP Week 14/Supermarket_Dataset_1 - Sales Data.csv")
head(sales)

```




# Check the Data
```{r Dimensions}
dim(sales)

```
```{r Structure}
str(sales)

```
```{r Categorical Var}
#Check their unique value counts
table(sales$Branch)
table(sales$Customer.type)
table(sales$Gender)
table(sales$Product.line)
table(sales$Payment)

#Convert to Factor
library(dplyr)

sales <- mutate_at(sales, vars(Branch, Customer.type,Gender,Product.line,Payment), as.factor)

str(sales)

```


# Data Cleaning
```{r Drop Irrelevant Columns}
# Droppig the Invoice Column
sales_new <- sales[,-1]
head(sales_new)

```
```{r Lower Case Columns}

# Lower Casing the Column Names
lower <- function(x) {
    colnames(x) <- tolower(colnames(x))
    x
}

sales_new <- lower(sales_new)

head(sales_new)

```

```{r Missing Data}
# check missing data
colSums(is.na(sales_new))

# No missing data
```
```{r Duplicated}
# Cheking for Duplicated records
sales_new <- unique(sales_new)
dim(sales_new)
```
```{r Outliers}
# Check for outliers
num_col <- sales_new[, c(5,6,7,11,12,13,14,15)]
outlier_detection = function(x){
  for(i in colnames(x)){
    boxplot(sales_new[[i]], xlab=i, main=paste0("Boxplot for ",i))
  }
}

outlier_detection(num_col)
```
```{r Dealing with Outliers}
# replace outliers with the 5th and 95th percentile
outlier_replace <- function(x){
   qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
   caps <- quantile(x, probs=c(.05, .95), na.rm = T)
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}


sales_new$tax <-outlier_replace(sales_new$tax)
sales_new$cogs <- outlier_replace(sales_new$cogs)
sales_new$gross.income <- outlier_replace(sales_new$gross.income)
sales_new$total <- outlier_replace(sales_new$total)

outlier_detection(num_col)


```
The Data had 1000 records and 16 columns
* 6 Categorical Variables
* 11 Numeric Variables

The Invoice.Id column was dropped because it was necessary for Analysis and all the columns were set to lower case for easier implementation
There are no missing values or duplicated values in our data set

Categorical Variables were converted to Factors which were:-

1. Branch
2. Customer Type
3. Gender
4. Product Line
5. Payment

In the Numeric variables there were outliers in:-

1. Tax
2. Total
3. Cogs
4. Gross_Income

Where the outliers were replaced by the 5th and 95th percentile


# Perform Exploratory Data Analysis
## Univariate
### Categorical Variables
```{r Categorical BarPLots}
library(ggplot2)

# Customer Type
customer_type <- ggplot(sales_new ,aes(x=customer.type , fill=customer.type)) + geom_bar() + labs(title = "Distribution of the Customer Types", x="Customer Type")
customer_type +scale_fill_discrete(name = "Customer Type") 

# How many Branches we have
bran <- ggplot(sales_new, aes(x=branch, fill=branch)) + geom_bar()+labs(title = "Distribution of Branches", x="Branches")+scale_fill_discrete(name = "Branch") 
bran

# Gender Distribution
gender1 <- ggplot(sales_new, aes(x=gender, fill=gender)) + geom_bar()+labs(title = "Distribution of Gender", x="Gender")+scale_fill_discrete(name = "Gender") 
gender1

# Types Product Line
product_line <- ggplot(sales_new,aes(x=product.line, fill=product.line))+ geom_bar()+ labs(title = "Distribution of Product Line", x="Product Line")+ scale_fill_discrete(name = "Product Line")
product_line + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Types Payments
payment_meth <- ggplot(sales_new, aes(x=payment, fill=payment)) + geom_bar()+labs(title = "Distribution of Payments", x="Payment")+scale_fill_discrete(name = "Payment") 
payment_meth

# Which Product line was bought more in terms of quantity per branch
product_line1 <- ggplot(sales_new,aes(x=product.line,y=quantity,fill=product.line))+geom_bar(stat = "identity")+ facet_wrap(~branch, scales = "free") + labs(title = "Total Quantity of each Product Line", subtitle = "[per Branch]",x="Product Line", y="Total Quantity") + scale_fill_discrete(name = "Product Line") + theme(axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())

product_line1

# Customer Type per Branch
customer_type1 <- ggplot(sales_new, aes(x=branch, fill=customer.type)) + geom_bar(position = "dodge")+labs(title = "Distribution of Customer Type per Branch", x="Payment")+scale_fill_discrete(name = " Customer Type") 
customer_type1

# Gender per Branch
gender2 <- ggplot(sales_new, aes(x=branch, fill=gender)) + geom_bar(position = "dodge")+labs(title = "Distribution of Gender per Branch", x="Branch")+scale_fill_discrete(name = "Gender") 
gender2

# Mode of Payment per Branch
payment_meth1 <- ggplot(sales_new, aes(x=branch, fill=payment)) + geom_bar(position = "dodge")+labs(title = "Distribution of Payment Mode per Branch", x="Branch")+scale_fill_discrete(name = "Payment") 
payment_meth1

# Customer Type and Mode of Payment
customer_type2 <- ggplot(sales_new, aes(x=customer.type, fill=payment)) + geom_bar(position = "dodge")+labs(title = "Customer Type and Mode of Payment", x="Customer Type")+scale_fill_discrete(name = "Mode of Payment") 
customer_type2

# Customer Type and Product Line
product_line2 <- ggplot(sales_new, aes(x=product.line, fill=customer.type)) + geom_bar(position = "dodge")+labs(title = "Distribution of Product Line for Each Customer Type", x="Product Line")+scale_fill_discrete(name = "Customer Type") 
product_line2 +theme(axis.text.x = element_text(angle = 30, hjust=0.8))

# Product Line vs Gender
product_line3 <- ggplot(sales_new, aes(x=product.line, fill=gender)) + geom_bar(position = "dodge")+labs(title = "Distribution of Product Line in Each Gender", x="Product Line")+scale_fill_discrete(name = "Gender") 
product_line3 +theme(axis.text.x = element_text(angle = 30, hjust=0.8))
 
```
### Numeric Variables
```{r Summary Stats, message=FALSE}
library(Hmisc)
describe(num_col)

```
```{r Histogram}

histogram = function(x){
  for(i in colnames(x)){
    hist(sales_new[[i]], breaks = 10,main =i,xlab = i,col = "dodgerblue")
  }
}

histogram(num_col)
```


## Bivariate
### Correlation
```{r Label Encoding}
# Label Encode Categorical Variables
encode_ordinal <- function(x, order = unique(x)) {
  x <- as.numeric(factor(x, levels = order, exclude = NULL))
  x
}
table(sales_new$branch, encode_ordinal(sales_new$branch))
table(sales_new$customer.type,encode_ordinal(sales_new$customer.type))
table(sales_new$gender, encode_ordinal(sales_new$gender))
table(sales_new$product.line, encode_ordinal(sales_new$product.line))
table(sales_new$payment, encode_ordinal(sales_new$payment))

sales_new$branch_enc <- encode_ordinal(sales_new$branch)
sales_new$customer.type_enc <- encode_ordinal(sales_new$customer.type)
sales_new$gender_enc <- encode_ordinal(sales_new$gender)
sales_new$product.line_enc <- encode_ordinal(sales_new$product.line)
sales_new$payment_enc <- encode_ordinal(sales_new$payment)

head(sales_new)
```

```{r Correlation}

# the correlation matrix
feat <- sales_new[, c(-1,-2,-3,-4,-8,-9,-10,-12)]
#head(feat)

correlation_matrix <- cor(feat)
library(corrplot)
corr_data <- as.data.frame(correlation_matrix)
corr_data

corr <- corrplot(correlation_matrix, method = "shade",order = "hclust", col=colorRampPalette(c("yellow","green","navyblue"))(100), addgrid.col="gray", tl.col = "black", tl.cex = 0.7,tl.srt = 45)
```

### Graph Analysis
```{r Categorical on Numeric, message=FALSE}
# Which Branch received the highest Average rating

by_rate <- sales_new %>%
  group_by(branch) %>%
  summarise(rate = mean(rating))

rate <- ggplot(by_rate, aes(x=branch, y= rate, fill=branch))+geom_bar(stat = "identity")+ labs(title = "Average Rating Per Branch", x="Branch",y="Average Rating")
rate + scale_fill_brewer(palette="Set1")+theme(legend.position = "none") 

# Product Line with Highest Gross Income
by_gross <- sales_new %>%
  group_by(product.line) %>%
  summarise(gross = sum(gross.income))


gross_inc <- ggplot(by_gross, aes(x=product.line, y= gross, fill=product.line))+geom_bar(stat = "identity")+ labs(title = "Product Line vs Gross Income", x="Product Line",y="Total Gross Income")
gross_inc + scale_fill_brewer(name="Product Line",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Total Sales for Each Product Line
by_sale <- sales_new %>%
  group_by(product.line) %>%
  summarise(total = sum(total))


total_sale <- ggplot(by_sale, aes(x=product.line, y= total, fill=product.line))+geom_bar(stat = "identity")+ labs(title = "Total Sales from Each Product Line", x="Product Line",y="Total Sales")
total_sale + scale_fill_brewer(name="Product Line",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Which Customer Type Generates more Gross Income
by_cust <- sales_new %>%
  group_by(customer.type) %>%
  summarise(gross = sum(gross.income))

cust_type <- ggplot(by_cust, aes(x=customer.type, y= gross, fill=customer.type))+geom_bar(stat = "identity")+ labs(title = " Total Gross Income per Customer Type", x="Customer Type",y="Total Gross Income")
cust_type + scale_fill_brewer( name= "Customer Type",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Branch with the Highest Gross Income
by_branch <- sales_new %>%
  group_by(branch) %>%
  summarise(gross = sum(gross.income))

branch_gross <- ggplot(by_branch, aes(x=branch, y= gross, fill=branch))+geom_bar(stat = "identity")+ labs(title = " Total Gross Income per Branch", x="Branch",y="Total Gross Income")
branch_gross + scale_fill_brewer(name="Branch",palette="Set1")+theme(
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Monthly Gross Income

sales_new$date <- as.Date(sales_new$date, format = "%m/%d/%Y")

by_date <- sales_new %>%
  group_by(date, branch) %>%
  summarise(gross = sum(gross.income))


month_gross <- ggplot(by_date, aes(x=date, y=gross)) +geom_line(color="blue") + labs(title = " Monthly Total Gross Income per Branch", x="",y="Total Gross Income")+facet_wrap(~branch, scales = "free_x", ncol=2)
month_gross


```
```{r Scatter Plot}
plot(sales_new$gross.income, sales_new$cogs, main = "Relationship between COGS and Gross Income", xlab="Gross Income", ylab="COGS")

plot(sales_new$total, sales_new$gross.income, main = "Total Sales vs. Gross Income ", xlab = "Total Sales", ylab = "Gross Income")

plot(sales_new$total, sales_new$rating, main = "Total Sales vs. Rating", xlab = "Total Sales", ylab = "Rating")
```



# Implement the Solution
## Dimension Reduction

```{r PCA}
# PCA

# Features
predictors <- sales_new[, c(-1,-2,-3,-4,-8,-9,-10,-12,-15)]

sales.pca <- prcomp(predictors, scale. = TRUE, center = TRUE)
summary(sales.pca)

label <- sales_new[, "total"]
head(label)

library(ggbiplot)
ggbiplot(sales.pca, obs.scale = 1.2,var.scale=1.3)
ggbiplot(sales.pca, choices = c(3,4) ,obs.scale = 1.2, var.scale = 1.3)


```

## Feature Selection
```{r Filter Method}
library(caret)

correlation = cor(predictors)

highly <- findCorrelation(correlation, cutoff = 0.75)
# highly correlated variables
names(predictors[,highly])

# Dataset without the Highly Correlated
sales_new1 <- predictors[-highly]

#Plot
par(mfrow = c(1, 2))
corrplot(correlation, order = "hclust",tl.col = "black", tl.cex = 0.7)
corrplot(cor(sales_new1), order = "hclust",tl.col = "black", tl.cex = 0.7)

```
```{r Wrapper Method}
library(clustvarsel)
library(mclust)

# perfectly correlated features

predictors_corr <- predictors[,c(-3,-4,-5)]

out = clustvarsel(predictors_corr, G=1:9)
out

# Features selected
data <- predictors_corr[,out$subset]
head(data)
```

```{r Ensemble Methods}
library("wskm")
library("cluster")

set.seed(2)
model <- ewkm(predictors, centers=5, lambda=5, maxiter=1000)

summary(model)

clusplot(predictors, model$cluster, color=TRUE, shade=TRUE,
         labels=5, lines=0,main='Cluster Analysis for Sales')

round(model$weights*100,2)

```

```{r Feature Ranking}
# Feature Ranking

# Removed Perfectly Correlated Variables
library(FSelector)

head(sales_new)

predictors1 <- sales_new[, c(-1,-2,-3,-4,-8,-9,-10,-11,-12,-13)]

Scores <- linear.correlation(total~., predictors1)
Scores

```

# Challenge the Solution
