---
title: "Marketing Strategies for Carrefour Kenya"
author: "Mercy M. Kubania"
date: "7/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Statement
> As a Data analyst at Carrefour Kenya I am doing a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). 

## a)Specifying the Question
1. Build models that will inform on relevant marketing strategies that will result in the highest no. of sales (total price including tax) for the Marketing Department at CarreFour Kenya

## b)The Metric of Success
> Analyse the data and Build a model that will inform on the relevant marketing strategies for the Marketing Department

## c)The Context
> Working on the CarreFour Kenya Sales data to identify the most relevant marketing strategies


## d)Experimental Design
1. Source the Data
2. Load the Data
3. Check the Data and Clean it
4. Perform Exploratory Data Analysis
5. Implement the Solution
6. Challenge the Solution

# Data Sourcing
```{r Dimension and Feature Data}
# Data for Analysis, Dimension Reduction and Feature Selection

sales <- read.csv("~/Documents/R markdowns/IP Week 14/Supermarket_Dataset_1 - Sales Data.csv")
head(sales)

```




# Check the Data
```{r Dimensions}
dim(sales)

```
```{r Structure}
str(sales)

```
```{r Categorical Var}
#Check their unique value counts
table(sales$Branch)
table(sales$Customer.type)
table(sales$Gender)
table(sales$Product.line)
table(sales$Payment)

#Convert to Factor
library(dplyr)

sales <- mutate_at(sales, vars(Branch, Customer.type,Gender,Product.line,Payment), as.factor)

str(sales)

```


# Data Cleaning
```{r Drop Irrelevant Columns}
# Droppig the Invoice Column
sales_new <- sales[,-1]
head(sales_new)

```
```{r Lower Case Columns}

# Lower Casing the Column Names
lower <- function(x) {
    colnames(x) <- tolower(colnames(x))
    x
}

sales_new <- lower(sales_new)

head(sales_new)

```

```{r Missing Data}
# check missing data
colSums(is.na(sales_new))

# No missing data
```
```{r Duplicated}
# Cheking for Duplicated records
sales_new <- unique(sales_new)
dim(sales_new)
```
```{r Outliers}
# Check for outliers
num_col <- sales_new[, c(5,6,7,11,12,13,14,15)]
outlier_detection = function(x){
  for(i in colnames(x)){
    boxplot(sales_new[[i]], xlab=i, main=paste0("Boxplot for ",i))
  }
}

outlier_detection(num_col)
```
```{r Dealing with Outliers}
# replace outliers with the 5th and 95th percentile
outlier_replace <- function(x){
   qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
   caps <- quantile(x, probs=c(.05, .95), na.rm = T)
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}


sales_new$tax <-outlier_replace(sales_new$tax)
sales_new$cogs <- outlier_replace(sales_new$cogs)
sales_new$gross.income <- outlier_replace(sales_new$gross.income)
sales_new$total <- outlier_replace(sales_new$total)

outlier_detection(num_col)


```
The Data had 1000 records and 16 columns

* 6 Categorical Variables
* 11 Numeric Variables

The Invoice.Id column was dropped because it was necessary for Analysis and all the columns were set to lower case for easier implementation
There are no missing values or duplicated values in our data set

Categorical Variables were converted to Factors which were:-

1. Branch
2. Customer Type
3. Gender
4. Product Line
5. Payment

In the Numeric variables there were outliers in:-

1. Tax
2. Total
3. Cogs
4. Gross_Income

Where the outliers were replaced by the 5th and 95th percentile


# Perform Exploratory Data Analysis
## Univariate
### Categorical Variables
```{r Categorical BarPLots}
library(ggplot2)

# Customer Type
customer_type <- ggplot(sales_new ,aes(x=customer.type , fill=customer.type)) + geom_bar() + labs(title = "Distribution of the Customer Types", x="Customer Type")
customer_type +scale_fill_discrete(name = "Customer Type") 

# How many Branches we have
bran <- ggplot(sales_new, aes(x=branch, fill=branch)) + geom_bar()+labs(title = "Distribution of Branches", x="Branches")+scale_fill_discrete(name = "Branch") 
bran

# Gender Distribution
gender1 <- ggplot(sales_new, aes(x=gender, fill=gender)) + geom_bar()+labs(title = "Distribution of Gender", x="Gender")+scale_fill_discrete(name = "Gender") 
gender1

# Types Product Line
product_line <- ggplot(sales_new,aes(x=product.line, fill=product.line))+ geom_bar()+ labs(title = "Distribution of Product Line", x="Product Line")+ scale_fill_discrete(name = "Product Line")
product_line + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Types Payments
payment_meth <- ggplot(sales_new, aes(x=payment, fill=payment)) + geom_bar()+labs(title = "Distribution of Payments", x="Payment")+scale_fill_discrete(name = "Payment") 
payment_meth

# Which Product line was bought more in terms of quantity per branch
product_line1 <- ggplot(sales_new,aes(x=product.line,y=quantity,fill=product.line))+geom_bar(stat = "identity")+ facet_wrap(~branch, scales = "free") + labs(title = "Total Quantity of each Product Line", subtitle = "[per Branch]",x="Product Line", y="Total Quantity") + scale_fill_discrete(name = "Product Line") + theme(axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())

product_line1

# Customer Type per Branch
customer_type1 <- ggplot(sales_new, aes(x=branch, fill=customer.type)) + geom_bar(position = "dodge")+labs(title = "Distribution of Customer Type per Branch", x="Payment")+scale_fill_discrete(name = " Customer Type") 
customer_type1

# Gender per Branch
gender2 <- ggplot(sales_new, aes(x=branch, fill=gender)) + geom_bar(position = "dodge")+labs(title = "Distribution of Gender per Branch", x="Branch")+scale_fill_discrete(name = "Gender") 
gender2

# Mode of Payment per Branch
payment_meth1 <- ggplot(sales_new, aes(x=branch, fill=payment)) + geom_bar(position = "dodge")+labs(title = "Distribution of Payment Mode per Branch", x="Branch")+scale_fill_discrete(name = "Payment") 
payment_meth1

# Customer Type and Mode of Payment
customer_type2 <- ggplot(sales_new, aes(x=customer.type, fill=payment)) + geom_bar(position = "dodge")+labs(title = "Customer Type and Mode of Payment", x="Customer Type")+scale_fill_discrete(name = "Mode of Payment") 
customer_type2

# Customer Type and Product Line
product_line2 <- ggplot(sales_new, aes(x=product.line, fill=customer.type)) + geom_bar(position = "dodge")+labs(title = "Distribution of Product Line for Each Customer Type", x="Product Line")+scale_fill_discrete(name = "Customer Type") 
product_line2 +theme(axis.text.x = element_text(angle = 30, hjust=0.8))

# Product Line vs Gender
product_line3 <- ggplot(sales_new, aes(x=product.line, fill=gender)) + geom_bar(position = "dodge")+labs(title = "Distribution of Product Line in Each Gender", x="Product Line")+scale_fill_discrete(name = "Gender") 
product_line3 +theme(axis.text.x = element_text(angle = 30, hjust=0.8))
 
```
### Numeric Variables
```{r Summary Stats, message=FALSE}
library(Hmisc)
describe(num_col)

```
```{r Histogram}

histogram = function(x){
  for(i in colnames(x)){
    hist(sales_new[[i]], breaks = 10,main =i,xlab = i,col = "dodgerblue")
  }
}

histogram(num_col)
```


## Bivariate
### Correlation
```{r Label Encoding}
# Label Encode Categorical Variables
encode_ordinal <- function(x, order = unique(x)) {
  x <- as.numeric(factor(x, levels = order, exclude = NULL))
  x
}
table(sales_new$branch, encode_ordinal(sales_new$branch))
table(sales_new$customer.type,encode_ordinal(sales_new$customer.type))
table(sales_new$gender, encode_ordinal(sales_new$gender))
table(sales_new$product.line, encode_ordinal(sales_new$product.line))
table(sales_new$payment, encode_ordinal(sales_new$payment))

sales_new$branch_enc <- encode_ordinal(sales_new$branch)
sales_new$customer.type_enc <- encode_ordinal(sales_new$customer.type)
sales_new$gender_enc <- encode_ordinal(sales_new$gender)
sales_new$product.line_enc <- encode_ordinal(sales_new$product.line)
sales_new$payment_enc <- encode_ordinal(sales_new$payment)

head(sales_new)
```

```{r Correlation}

# the correlation matrix
feat <- sales_new[, c(-1,-2,-3,-4,-8,-9,-10,-12)]
#head(feat)

correlation_matrix <- cor(feat)
library(corrplot)
corr_data <- as.data.frame(correlation_matrix)
corr_data

corr <- corrplot(correlation_matrix, method = "shade",order = "hclust", col=colorRampPalette(c("yellow","green","navyblue"))(100), addgrid.col="gray", tl.col = "black", tl.cex = 0.7,tl.srt = 45)
```

### Graph Analysis
```{r Categorical on Numeric, message=FALSE}
# Which Branch received the highest Average rating

by_rate <- sales_new %>%
  group_by(branch) %>%
  summarise(rate = mean(rating))

rate <- ggplot(by_rate, aes(x=branch, y= rate, fill=branch))+geom_bar(stat = "identity")+ labs(title = "Average Rating Per Branch", x="Branch",y="Average Rating")
rate + scale_fill_brewer(palette="Set1")+theme(legend.position = "none") 

# Product Line with Highest Gross Income
by_gross <- sales_new %>%
  group_by(product.line) %>%
  summarise(gross = sum(gross.income))


gross_inc <- ggplot(by_gross, aes(x=product.line, y= gross, fill=product.line))+geom_bar(stat = "identity")+ labs(title = "Product Line vs Gross Income", x="Product Line",y="Total Gross Income")
gross_inc + scale_fill_brewer(name="Product Line",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Total Sales for Each Product Line
by_sale <- sales_new %>%
  group_by(product.line) %>%
  summarise(total = sum(total))


total_sale <- ggplot(by_sale, aes(x=product.line, y= total, fill=product.line))+geom_bar(stat = "identity")+ labs(title = "Total Sales from Each Product Line", x="Product Line",y="Total Sales")
total_sale + scale_fill_brewer(name="Product Line",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Which Customer Type Generates more Gross Income
by_cust <- sales_new %>%
  group_by(customer.type) %>%
  summarise(gross = sum(gross.income))

cust_type <- ggplot(by_cust, aes(x=customer.type, y= gross, fill=customer.type))+geom_bar(stat = "identity")+ labs(title = " Total Gross Income per Customer Type", x="Customer Type",y="Total Gross Income")
cust_type + scale_fill_brewer( name= "Customer Type",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Branch with the Highest Gross Income
by_branch <- sales_new %>%
  group_by(branch) %>%
  summarise(gross = sum(gross.income))

branch_gross <- ggplot(by_branch, aes(x=branch, y= gross, fill=branch))+geom_bar(stat = "identity")+ labs(title = " Total Gross Income per Branch", x="Branch",y="Total Gross Income")
branch_gross + scale_fill_brewer(name="Branch",palette="Set1")+theme(
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Monthly Gross Income

sales_new$date <- as.Date(sales_new$date, format = "%m/%d/%Y")

by_date <- sales_new %>%
  group_by(date, branch) %>%
  summarise(gross = sum(gross.income))


month_gross <- ggplot(by_date, aes(x=date, y=gross)) +geom_line(color="blue") + labs(title = " Monthly Total Gross Income per Branch", x="",y="Total Gross Income")+facet_wrap(~branch, scales = "free_x", ncol=2)
month_gross


```
```{r Scatter Plot}
plot(sales_new$gross.income, sales_new$cogs, main = "Relationship between COGS and Gross Income", xlab="Gross Income", ylab="COGS")

plot(sales_new$total, sales_new$gross.income, main = "Total Sales vs. Gross Income ", xlab = "Total Sales", ylab = "Gross Income")

plot(sales_new$total, sales_new$rating, main = "Total Sales vs. Rating", xlab = "Total Sales", ylab = "Rating")
```

In the Univariate Analysis we observed the following:-

Customer Type were only 2 types, and the both had equal distribution as well as the Gender
The most popular Branch is Branch A, B then C but the difference between the two is very minimal.
In Product Line, the *Fashion Accessories* were the most popular product line, followed by *Food & Beverages*, *Electronic Accessories* 
The most popular mode of Payment is through the E-Wallet

In the Branches there were also interesting observations, in Branch A the most popular Product Line are:-

1. Home & Lifestyle
2. Sports & Travel
3. Electronic Accessories

In Branch B we have:

1. Sports & Travel and Health & Beauty
2. Electronic Accessories

In Branch C we have:

1. Food & Beverages
2. Fashion & Accessories
3. Electronic Accessories

In terms of Most frequent Customer Types in the Branches we have

Branch A -> Has more Normal Customers than Members \n
Branch B -> Has more Normal Customers than Members, but the difference is slight \n
Branch C -> Has more Members than Normal Customers

In terms of the gender we see that 

Branch A -> Has more Males than Female Customers \n
Branch B -> Has more Males than Female Customers \n
Branch C -> Has more Females than Male Customers

This observations gives us interesting insights, we have the most popular Product Line in Branch A and B is mostly Sports & Travel and Electronics Accessories while the most frequent customers in this Branches are Male Customers and the most frequent Customer Type in this Branches are Normal Customers. 

We can say that in Branch A and B we have more Male Customers who are not members, Marketing team can look into targeting these customers in making them Members to establish a permanent relationship and also look into targeting Female Customers in this branches

In Branches C, we have more Female Customers with most bought products being Food & Beverages and Fashion Accessories, the most popular Customer Types are Members, therefore the Marketing Team can look into what influences the Female Customers in Branch C and apply it to Branch A and b
The Team can do the same for Branch C, look into what influences customers in Branch A and B and apply it.
Further Analysis can be done on the specific Braches to see what insights we can draw from their customer behaviors

In Branch A the most popular mode of Payment is Ewallet then cash. In Branch B, the most of Payment is EWallet, In Branch C we have Cash as the mode Payment.
Members Mostly use Credit Card and Normal Customers use EWallet. 

In Histogram there are no numeric columns are normally distributed.

The Categorical variables were encoded using Ordinal Encoding and created new columns.

In the Correlation we see that the features:-

1. Cogs
2. Total
3. Gross Income

were perfectly correlated with a correlation of 1

In the assumption that rating was done for Branch, we see that Branch C has the highest average Rating
The Product Line with the Highest Gross Income:-

1. Food & Beverages
2. Sports & Travels
3. Electronic Accessories

The Marketing and Sales can look into the Product Lines so as to develop marketing strategies for those other Product Lines

We see that in between the Months Beginning of January in all Branches have low Total Gross Income but Mid January towards the Beginning of February there is relatively Highly Gross Income.

At the end of  March the Total Gross Income lowers except for Branch B where it starts to pick up

In Branch B between End of January and Beginning of March there is relatively High Total Gross Income


# Implement the Solution
## Dimension Reduction

```{r PCA}
# PCA

# Features
predictors <- sales_new[, c(-1,-2,-3,-4,-8,-9,-10,-12,-15)]

sales.pca <- prcomp(predictors, scale. = TRUE, center = TRUE)
summary(sales.pca)

label <- sales_new[, "total"]

library(ggbiplot)
ggbiplot(sales.pca, obs.scale = 1.2,var.scale=1.3)
ggbiplot(sales.pca, choices = c(3,4) ,obs.scale = 1.2, var.scale = 1.3)


```

## Feature Selection
```{r Filter Method}
library(caret)

correlation = cor(predictors)

highly <- findCorrelation(correlation, cutoff = 0.75)
# highly correlated variables
names(predictors[,highly])

# Dataset without the Highly Correlated
sales_new1 <- predictors[-highly]

#Plot
par(mfrow = c(1, 2))
corrplot(correlation, order = "hclust",tl.col = "black", tl.cex = 0.7)
corrplot(cor(sales_new1), order = "hclust",tl.col = "black", tl.cex = 0.7)

```
```{r Wrapper Method}
library(clustvarsel)
library(mclust)

# perfectly correlated features

predictors_corr <- predictors[,c(-3,-4,-5)]

out = clustvarsel(predictors_corr, G=1:9)
out

# Features selected
data <- predictors_corr[,out$subset]
head(data)
```


```{r Feature Ranking}
# Feature Ranking

# Removed Perfectly Correlated Variables
library(FSelector)

head(sales_new)

predictors1 <- sales_new[, c(-1,-2,-3,-4,-7,-8,-9,-10,-11,-12,-13)]

Scores <- linear.correlation(total~., predictors1)
Scores

# Selected Features
Subset <- cutoff.k(Scores, 5)
as.data.frame(Subset)

```


The Problem we are tackling is Regression problem with the *label* being the *Total Column* and for Dimension Reduction Principal Component Analysis was used and there were 11 Principal Components generated and PC1 to PC6 are able to explain 84% of the variations in the data.

The Categorical Variables were encoded through Ordinal Encoding and new columns were created.
In Feature Selection the different methods were used.
In Filter Methods with a cutoff at 0.7, the variables accepted were:

1. Branch
2. Customer Type
3. Gender
4. Product Line
5. Payment
6. Unit Price
7. Quantity
8. Gross Income
9. Rating


Excluding the Total column for it the Label column

The Wrapper method selected, it excluded the perfectly correlated features and it returned

1. Quantity
2. Payment

The Feature Ranking while excluding the perfectly correlated features were:

1. Quantity
2. Unit Price
3. Gender
4. Product Line
5. Rating


When building an Unsupervised or Supervised model to predict on Total Sales the Features generated by each of the methods could be used and see which group of Features will generate the best accuracy
